---
title: "Evaluation of Sound Reduction Index R Comparison between Maximum Likelihood and Bayes Statistics"
author: "Michael Parzinger"
date: "`r Sys.Date()`"
output: 
  html_document:
    number_sections: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F)

realDataAvailable <- TRUE

library(SoundModeling) # https://github.com/Michi-Parz/SoundModeling
if (realDataAvailable)
  library(SoundModelingData)
library(PupillometryR)
library(ggridges)

numb_curves <- nrow(sldf)/21

#colorsUsed <- RavenR::ggplotColours(4)



col_fre <- "#0066cc"
col_bay <- "#d62728"
col_mean <- "#000000"

col_vec <- c(col_bay, col_fre,col_mean)


```


# Important informations


In this project, you must add both a folder named *Bayes* and a folder named *ML*.


The folder *ML* must contain the projekt https://github.com/Michi-Parz/SRI-Model-MLE

The folder *Bayes* must contain the projekt https://github.com/Michi-Parz/SRI-Model-Bayes



Otherwise, you will need to update the file paths listed in this document!


```{r}
# Lade Bayes Schätzungen

poster <- readRDS("Bayes/Saves/tvAR/STAN_FixMaxRL_tvAR_max_shift_1")
exposter <- rstan::extract(poster)
```


```{r}
# Erzeuge Freq Model
formu_final <- RpdB ~ soundproofing_measure(
    eta_int, C, cL, 
    w_perimeter = u,
    w_surface = s,
    freq = fpHz,
    fce_gamma = gamma,
    cL1 = 2500,
    plateau_style = "tau0real",
    eta_ref = NULL,
    max_rl = 1.18,
    max_rl_type = "value"
  )


f_model <- nlsLMcov(formu_final,
                        sldf,
                        start = c("eta_int" = 0.02,"C" = 2,"cL" = 1000, "gamma" = 0.77),
                        cov_type = "tvAR",
                        lower = c(5/1000, 0,0,0),
                        maxlag = 1L,
                        )

```

```{r}
# Lade Bootstrapp Ziehungen
bs_para_file <- paste("ML/Saves/BS_para_tvAR", 1,"_maxrlvalue_SpaceSaved", sep = "")

bs_fits <- readRDS(bs_para_file)

bs_coefs <- t(sapply(bs_fits, coefficients))
bs_coefs <- as.data.frame(bs_coefs)
```

```{r}
## Bootstrapp Residuen
# 
bs_resid <- lapply(bs_fits, \(x){
  matrix(sldf$RpdB  - c(x$fitted), ncol = 21, byrow = T)
  
  #matrix(x$resid, ncol = 21, byrow = T)
})
bs_resid <- abind::abind(bs_resid, rev.along = 3)
# Standardisiert
bs_st_resid <- lapply(bs_fits, \(x){
  cov <- x$cov_matrix
  weight <- solve(chol(cov))

  res <- (sldf$RpdB  - c(x$fitted)) %*% kronecker(diag(numb_curves),weight)
  matrix(res, ncol = 21, byrow = T)
  
  #matrix(resid(x), ncol = 21, byrow = T)
})
bs_st_resid <- abind::abind(bs_st_resid, rev.along = 3)
```




# Model curves

```{r}

sldf_fit <- sldf

sldf_fit$Mean <- rep(apply(sl_list_stan$y, 2, mean), numb_curves)
sldf_fit$Frequentist <- f_model$fitted
sldf_fit$Bayes <- c(t(apply(exposter$SRI, 2:3, mean)))

ggModelCurves <- gg_basic(sldf_fit, aes(x = fpHz, group = Building), ymin = 20, ymax = 80)+
  geom_point(aes(y = RpdB), color = "grey")+
  geom_line(aes(y = RpdB), color = "grey", alpha = 1/3)+
  geom_line(aes(y = Frequentist, color = "Max likelihood"))+
  geom_line(aes(y = Bayes, color = "Bayesian"))+
  geom_line(aes(y = Mean, color = "Data mean"))+
  scale_color_manual(values = col_vec,
                     breaks = c("Bayesian", "Max likelihood", "Data mean")
                       )+
  labs(color = "")+
  theme(
    legend.position = c(0.7, 0.2),
    legend.justification = c(0.5, 0.5),
    legend.background = element_rect(fill = "transparent"),
    legend.spacing.y = unit(-0.7, "cm"),
    panel.grid.major = element_line(linewidth = 1))

ggModelCurves

```


# Mahalanobis


```{r}
## Bayes
# Alle
sqmaha_dist_all_b <- apply(exposter$st_resid, 1, \(x){
    t(c(x))%*%c(x)
})


sqmaha_dist_build_b <- apply(
    exposter$st_resid^2,
    1:2, sum
  )


sqmaha_dist_build_b <- tidyr::pivot_longer(data.frame(sqmaha_dist_build_b), 1:27,
                         names_to = "Building",
                         values_to = "SQMahalanobis")

sqmaha_dist_build_b$Building <- ordered(rep(1:27, 4000), 1:27)
sqmaha_dist_build_b$Method <- "Bayesian"

```

```{r}
# ML
sqmaha_dist_all_c <- apply(bs_st_resid, 1, \(x){
    t(c(x))%*%c(x)
})



sqmaha_dist_build_c <- apply(
    bs_st_resid^2,
    1:2, sum
  )

sqmaha_dist_build_c <- tidyr::pivot_longer(data.frame(sqmaha_dist_build_c), 1:27,
                         names_to = "Building",
                         values_to = "SQMahalanobis")


sqmaha_dist_build_c$Building <- ordered(rep(1:27, 1000), 1:27)
sqmaha_dist_build_c$Method <- "Frequentist"


```


```{r}
maha_all_df <- data.frame(
  sq_maha = c(sqmaha_dist_all_b, sqmaha_dist_all_c, rchisq(3000, 21*numb_curves)),
  Method = c(rep("Bayesian", 4000), rep("Frequentist", 1000), rep("Chi", 3000))
)


maha_all_df$Method <- factor(maha_all_df$Method, 
                    levels = c("Bayesian", "Frequentist", "Chi"))

ggplot(maha_all_df)+
  aes(x = Method, y = sqrt(sq_maha), fill = Method)+
  geom_boxplot()+
  scale_fill_manual(breaks = c("Bayesian", "Frequentist", "Chi"), values = c(col_vec, "grey"))+
  labs(y = "Mahalanobis distance")+
  theme(legend.position = "none")



```
```{r}
# sqmaha_chi <- data.frame(
#   "Building" = 0,
#   "SQMahalanobis" = rchisq(3000,21),
#   "Method" = "Chi"
# )

sqmaha_dist_build <- rbind(sqmaha_dist_build_b, sqmaha_dist_build_c)#,sqmaha_chi)
# sqmaha_dist_build$Method <- factor(sqmaha_dist_build$Method, 
#                     levels = c("Bayesian", "Frequentist", "Chi"))

ggplot(sqmaha_dist_build)+
  aes(x = Building, y = sqrt(SQMahalanobis ), fill = Method)+
  geom_boxplot()+
  scale_fill_manual(values = col_vec)+
  labs(y = "Mahalanobis distance")+
  theme(legend.position = "bottom")

```

```{r}
ggplot(sqmaha_dist_build) +
  aes(x = sqrt(SQMahalanobis), y = Building, fill = Method)+
    geom_density_ridges2(alpha = 1/3) +
    theme(axis.text.x = element_text(size = 15),
          axis.title = element_text(size = 15),
          panel.grid.major = element_line(linewidth = 1))+
  labs(x = "Mahalanobis distance",
       y = "Building")+
  theme(legend.position = "bottom")+
  scale_fill_manual(values = col_vec)
```



# Fit parameter

```{r}

bs_coefs$Method <- "Bootstrapping"

mcmc_coefs <- data.frame(exposter[1:4])

mcmc_coefs$Method <- "MCMC"

sample_coef_df <- rbind(bs_coefs,mcmc_coefs )



```

```{r}


sample_coef_df$Method <- factor(sample_coef_df$Method,
                                levels = c("MCMC","Bootstrapping"))

gg_kernel <- ggplot(sample_coef_df, aes(x= Method, fill = Method))

# eta_int
gg_ei <- gg_kernel+
  aes(y=eta_int)+
  geom_flat_violin(position = position_nudge(x = 0),
                   alpha = 0.7) +
  geom_boxplot(width = 0.1) +
  labs(y = bquote(eta["int"]))+
  theme(legend.position = "none")+
  scale_fill_manual(values = col_vec,
                    breaks = c("MCMC", "Bootstrapping"))+
  coord_cartesian(ylim = c(5/1000,0.045))

# C
gg_C <- gg_kernel+
  aes(y=C)+
  geom_flat_violin(position = position_nudge(x = 0),
                   alpha = 0.7) +
  geom_boxplot(width = 0.1) +
  labs(y = "C")+
  theme(legend.position = "none")+
  scale_fill_manual(values = col_vec,
                    breaks = c("MCMC", "Bootstrapping"))

# cL
gg_cL <- gg_kernel + aes(y=cL)+
  geom_flat_violin(position = position_nudge(x = 0),
                   alpha = 0.7) +
  geom_boxplot(width = 0.1) +
  coord_cartesian(ylim = c(900,1150))+
  labs(y = "cL / m/s")+
  theme(legend.position = "none") +
  scale_fill_manual(values = col_vec,
                    breaks = c("MCMC", "Bootstrapping"))

# gamma
gg_gamma <- gg_kernel + aes(y=gamma)+
  geom_flat_violin(position = position_nudge(x = 0),
                   alpha = 0.7) +
  geom_boxplot(width = 0.1) +
  labs(y = bquote(gamma))+
  theme(legend.position = "none") +
  scale_fill_manual(values = col_vec,
                    breaks = c("MCMC", "Bootstrapping"))

# Alle

ggarrange(gg_ei, gg_C, gg_cL, gg_gamma)


```


# Standard deviation

```{r}
quantiles <- c(0.025, 0.5, 0.975)
```


```{r}
# Classic
stdev_classic <- sapply(bs_fits, \(x){
  cov <- x$cov_matrix
  sqrt(diag(cov))
})
stdev_classic <- t(stdev_classic)

stdev_c_quant <- apply(stdev_classic, 2, quantile, probs = quantiles)

stdev_c_df <- data.frame(
  t(stdev_c_quant),
  fpHz = freq,
  Method = "Bootstrapping"
)
names(stdev_c_df)[1:3] <- paste("q", c(2.5,50,97.5), sep = "")



# Bayes
# var_list_bayes <- apply(exposter$cov_mat, 1, diag, simplify = F)
# stdev_bayes <- sapply(var_list_bayes, sqrt)
# stdev_bayes <- t(stdev_bayes)

stdev_bayes <- exposter$unconditional_st_dev


stdev_b_quant <- apply(stdev_bayes, 2, quantile, probs = quantiles)

stdev_b_df <- data.frame(
  t(stdev_b_quant),
  fpHz = freq,
  Method = "MCMC"
)
names(stdev_b_df)[1:3] <- paste("q", c(2.5,50,97.5), sep = "")


# Kombiniert
stdev_df <- rbind(stdev_b_df, stdev_c_df)
```


```{r}

stdev_df$Method <- factor(stdev_df$Method,
                                levels = c("MCMC","Bootstrapping"))

ggplot(stdev_df)+
  aes(x = fpHz, group = Method, fill = Method, color = Method)+
  geom_line(aes(y = q50))+
  geom_ribbon(aes(ymin = q2.5, ymax = q97.5), alpha = 1/4)+
  scale_x_continuous(trans = log2_trans(),
                       minor_breaks = c(freq,6300,8000,10000),
                       breaks = c(63,125,250,500,1000,2000,4000,8000),
                       labels = c(63,125,250,500,1000,2000,4000,8000),
    )+
  labs(y = "Standard deviation / dB",
       x = "Frequency f / Hz")+
  theme(legend.position = "bottom")+
  scale_fill_manual(values = col_vec)


```
```{r}
# Classic data.frame
stdev_classic_df <- data.frame(stdev_classic)
names(stdev_classic_df) <- freq
stdev_classic_df_long <- tidyr::pivot_longer(stdev_classic_df,
                                           1:21,
                                           names_to = "fpHz",
                                           values_to = "StDev")


stdev_classic_df_long$Method <- "Frequentist"

# Bayes data.frame
stdev_bayes_df <- data.frame(stdev_bayes)
names(stdev_bayes_df) <- freq
stdev_bayes_df_long <- tidyr::pivot_longer(stdev_bayes_df,
                                           1:21,
                                           names_to = "fpHz",
                                           values_to = "StDev")
stdev_bayes_df_long$Method <- "Bayesian"
# Combine


stdev_water_df <- rbind(stdev_bayes_df_long,stdev_classic_df_long)

stdev_water_df$fpHz <- ordered(stdev_water_df$fpHz, as.factor(freq))

```

```{r}
ggplot(stdev_water_df) +
  aes(x = log2(StDev), y = fpHz, fill = Method)+
    geom_density_ridges2(alpha = 1/3) +
    theme(axis.text.x = element_text(size = 15),
          axis.title = element_text(size = 15),
          panel.grid.major = element_line(linewidth = 1))+
  labs(x = bquote(log[2]~standard~deviation),
       y = "Frequency f / Hz")+
  theme(legend.position = "bottom")+
  scale_fill_manual(values = col_vec)
```


# AR coefficients




```{r}
# Classic
arcoff_classic <- sapply(bs_fits, \(x){
  cov <- x$cov_matrix
  ar_mat <- tvARcov2coeff(cov)
  res <- c()
  for (i in 1:20) {
    res[i] <- ar_mat[i+1,i]
  }
  res
})
arcoff_classic <- t(arcoff_classic)

c_quant <- apply(arcoff_classic, 2, quantile, probs = quantiles)

ar_c_df <- data.frame(
  t(c_quant),
  fpHz = freq[-1],
  Method = "Bootstrapping"
)
names(ar_c_df)[1:3] <- paste("q", quantiles*100, sep = "")



# Bayes
arcoeff_bayes <- exposter$shift_coeff


b_quant <- apply(arcoeff_bayes, 2, quantile, probs = quantiles)

ar_b_df <- data.frame(
  t(b_quant),
  fpHz = freq[-1],
  Method = "MCMC"
)
names(ar_b_df)[1:3] <- paste("q", quantiles*100, sep = "")


# Kombiniert
ar_df <- rbind(ar_b_df, ar_c_df)
```


```{r}


ar_df$Method <- factor(ar_df$Method,
                                levels = c("MCMC","Bootstrapping"))

ggplot(ar_df)+
  aes(x = fpHz, group = Method, fill = Method, color = Method)+
  geom_line(aes(y = q50))+
  geom_ribbon(aes(ymin = q2.5, ymax = q97.5), alpha = 1/4)+
  scale_x_continuous(trans = log2_trans(),
                       minor_breaks = c(freq,6300,8000,10000),
                       breaks = c(63,125,250,500,1000,2000,4000,8000),
                       labels = c(63,125,250,500,1000,2000,4000,8000),
    )+
  labs(y = "AR coefficient",
       x = "Frequency f / Hz")+
  theme(legend.position = "bottom")+
  scale_fill_manual(values = col_vec)


```


Waterfall:

```{r}
# Classic data.frame
ar_classic_df <- data.frame(arcoff_classic)
names(ar_classic_df) <- freq[-1]
ar_classic_df <- tidyr::pivot_longer(ar_classic_df,
                                           1:20,
                                           names_to = "fpHz",
                                           values_to = "AR")


ar_classic_df$Method <- "Frequentist"

# Bayes data.frame
ar_bayes_df <- data.frame(arcoeff_bayes)
names(ar_bayes_df) <- freq[-1]
ar_bayes_df <- tidyr::pivot_longer(ar_bayes_df,
                                           1:20,
                                           names_to = "fpHz",
                                           values_to = "AR")
ar_bayes_df$Method <- "Bayesian"
# Combine


ar_water_df <- rbind(ar_bayes_df,ar_classic_df)

ar_water_df$fpHz <- ordered(ar_water_df$fpHz, as.factor(freq[-1]))

```

```{r}
ggplot(ar_water_df) +
  aes(x = AR, y = fpHz, fill = Method)+
    geom_density_ridges2(alpha = 1/3) +
    theme(axis.text.x = element_text(size = 15),
          axis.title = element_text(size = 15),
          panel.grid.major = element_line(linewidth = 1))+
  labs(x = "AR coefficient",
       y = "Frequency f / Hz")+
  theme(legend.position = "bottom")+
  scale_fill_manual(values = col_vec)
```



# Correlation

```{r}

corr_freq <- cov2cor(f_model$cov[[3]])
corr_bayes <- apply(exposter$cov_mat, 1, cov2cor, simplify = F)
corr_bayes <- abind(corr_bayes, rev.along = 3)
mean_corr_b <- apply(corr_bayes, 2:3, mean)


corr_comp <- corr_freq*lower.tri(diag(21), diag = TRUE) + mean_corr_b*t(lower.tri(diag(21)))
colnames(corr_comp) <- row.names(corr_comp) <- paste("f", freq)



```

```{r}
corrplot::corrplot(corr_comp, "color", tl.col = "black",
                   title = "L = Frequentist & U = Bayesian",
                   mar = c(0,0,1,0))

corr_diff <- corr_freq - mean_corr_b
colnames(corr_diff) <- row.names(corr_diff) <- paste("f", freq)

corrplot::corrplot(corr_diff, tl.col = "black",
                   is.corr = F,
                   title = "Frequentist - Bayesian",
                   mar = c(0,0,1,0))
```



```{r}

alpha <- 0.05

corr_lower <- apply(corr_bayes, 2:3, quantile, alpha/2)
corr_upper <- apply(corr_bayes, 2:3, quantile, 1-alpha/2)

cor_lower <- cov2cor(corr_lower)*lower.tri(diag(21), diag = TRUE)
cor_upper <- cov2cor(corr_upper)*lower.tri(diag(21))

cor_range <- cor_lower+t(cor_upper)

colnames(cor_range) <- row.names(cor_range) <- paste("f", freq)




```



```{r}

bs_cov_mats <- lapply(bs_fits, \(x){
  x$cov_matrix
})

bs_corr_mats <- lapply(bs_cov_mats, cov2cor)

bs_corr_array <- abind::abind(bs_corr_mats, along = 0)

alpha <- 0.05

bs_corr_min <- apply(bs_corr_array, 2:3, quantile, alpha/2)*lower.tri(diag(21))
bs_corr_max <- apply(bs_corr_array, 2:3, quantile, 1-alpha/2)*lower.tri(diag(21))

bs_corr_range <- bs_corr_min+t(bs_corr_max)+diag(21)

colnames(bs_corr_range) <- row.names(bs_corr_range) <- paste("f", freq)





```

```{r}

#par(mfrow = c(1,2))


corrplot::corrplot(
  cor_range,
  "color", tl.col = "black",
  title = bquote("Bayesian range: L = 2.5 % & U = 97.5 %"),
  mar = c(0,0,1,0)
)


corrplot::corrplot(
  bs_corr_range,
  "color", tl.col = "black",
  title = bquote("Frequentist range: L = 2.5 % & U = 97.5 %"),
  mar = c(0,0,1,0)
)



#par(mfrow = c(1,1))

```

# Rw Boxplots


```{r}
curves_bs_wall <- readRDS("ML/Saves/BS_sample_wall_thickness")
rw_values_c <- apply(curves_bs_wall, 1:2, rw_value)
rw_values_b <- t(readRDS("Bayes/Saves/RW_wall_samples"))
```


```{r}
# Adjust frequentist
rw_df_c <- data.frame(rw_values_c)
rw_df_c <- tidyr::pivot_longer(rw_df_c,cols = names(rw_df_c),
                                names_to = "Thickness", values_to = "Rw")
rw_df_c$Thickness <- rep(c(18:36)/100, 1000)

rw_df_c$Method <- "Bootstrapping"
```

```{r}
# Adjust bayes
rw_df_b <- data.frame(rw_values_b)
rw_df_b <- tidyr::pivot_longer(rw_df_b,cols = names(rw_df_b),
                                names_to = "Thickness", values_to = "Rw")
rw_df_b$Thickness <- rep(c(18:36)/100, 4000)

rw_df_b$Method <- "MCMC"
```


```{r}
rw_df <- rbind(rw_df_c, rw_df_b)
```

```{r}

rw_df$Method <- factor(rw_df$Method,
                                levels = c("MCMC","Bootstrapping"))

ggplot(rw_df)+
  aes(x = factor(Thickness), y = Rw, fill = Method)+
  geom_boxplot()+
  labs(y = bquote(Weighted~sound~reduction~index~R[w]),
       x = "Thickness in m")+
  scale_fill_manual(values = col_vec)+
  scale_y_continuous(
    minor_breaks = seq(30,80,1),
    breaks = seq(45,65,5),
    labels = seq(45,65,5)
  )+
  theme(legend.position = "bottom",
        panel.grid.major = element_line(linewidth  = 1))

```





# Rw probabilities



```{r}
rw_prob_b <- readRDS("Bayes/Saves/Rw_prop_df")
rw_prob_c <- readRDS("ML/Saves/Rw_prop_df")

rw_prob_b$Method <- "MCMC"
rw_prob_c$Method <- "Bootstrapping"

rw_prob <- rbind(rw_prob_b, rw_prob_c)
rw_prob <- rw_prob[rw_prob$Rw >47 & rw_prob$Rw < 59,]
rw_prob$Rw <- as.factor(rw_prob$Rw)
```





```{r}
rw_prob$Method <- factor(rw_prob$Method,
                                levels = c("MCMC","Bootstrapping"))

gg_rw <- ggplot(rw_prob)+
  aes(x = t, y = Prob,
      linetype = Rw, shape = Rw,
      color = Method,
      group = interaction(Method, Rw))+
  geom_line()+
  geom_point()+
  labs(y= bquote(Probability~of~italic(R)[w]>=~threshold),
       color = "Method",
       shape = "Threshold in dB",
       linetype = "Threshold in dB",
       x = "Thickness in m")+
    scale_x_continuous(
                       minor_breaks = (18:36)/100,
                       breaks = seq(0.18,0.36,0.04),
                       labels = seq(0.18,0.36,0.04))+
    scale_y_continuous(
        minor_breaks = c(0:100)/100,
        breaks = seq(0,1, 0.05)
    )+
  theme(axis.text = element_text(size = 15),
          axis.title = element_text(size = 15),
          plot.title = element_text(size = 15),
          legend.text = element_text(size = 15),
          panel.grid.major = element_line(linewidth = 1),
          axis.text.x = element_text(angle =0, hjust = 1),
        legend.position = "bottom")+
  scale_color_manual(values = col_vec)+
  scale_linetype_manual(values = c("dotted", "solid", "dashed"))+
  scale_shape_manual(values = c(15,16,17))

gg_rw
gg_rw+
  coord_cartesian(ylim = c(0.75, 1))

```




# Bonus

## Fit vs Residuals


```{r}

whitened_res_df <- data.frame(
  "fpHz" = rep(freq, 27),
  "fres" = f_model$m$resid(),
  "bres" = c(t(apply(exposter$st_resid,2:3, mean))),
  "ffit" = f_model$fitted,
  "bfit" = c(t(apply(exposter$SRI,2:3, mean)))
)


fitvsres <- ggplot(whitened_res_df)+
  geom_point(aes(x = ffit, y = fres, color = "Frequentist"))+
  geom_point(aes(x = bfit, y = bres, color = "Bayesian"))+
  labs(y = "Whitened residuals", x = "Model fit", color = "")+
  scale_color_manual(values = col_vec)



freqvsres <- gg_basic(whitened_res_df, aes(x = fpHz), sound = F, ymin = -3, ymax = 3)+
  geom_point(aes( y = fres, color = "Frequentist"))+
  geom_point(aes( y = bres, color = "Bayesian"))+
  labs(y = "Whitened residuals", color = "")+
  scale_color_manual(values = col_vec)
  


ggarrange(fitvsres, freqvsres, legend = "bottom", common.legend = T)



```




## Frequency vs Residuals (Boxplot)




```{r}

white_res_box <- data.frame(
  "fpHz" = rep(whitened_res_df$fpHz, 2),
  "Res" = c(whitened_res_df$fres, whitened_res_df$bres),
  "Method" = rep(c("Max likelihood", "Bayesian"), each = 567)
)



ggWR <- ggWhiteRes(white_res_box, aes(x = as.factor(fpHz),
                                      y = Res, fill = Method),
                   numbVariants = 2, tikz = TRUE, txtsize = 3.25)+
  labs(y = "Whitened residuals", x = expression(paste("Frequency ",
                             italic(" f"), " / Hz")),
       fill = "")+
  # scale_x_discrete( minor_breaks = c(freq,6300,8000,10000),
  #                      breaks = c(63,125,250,500,1000,2000,4000,8000),
  #                      labels = c(63,125,250,500,1000,2000,4000,8000),
  #   )+
  scale_fill_manual(values = col_vec)+
  theme(legend.position = "bottom",
        axis.text.x = element_text(angle =40, hjust = 1))


ggWR

```


## PPC One-Third Ocatave Band



```{r}

alphas <- c(0.5,0.05,0.01)

yRepQ <- apply(exposter$y_rep, 2:3, quantile,
               probs = c(alphas/2, 1-alphas/2))

yRepQdf <- data.frame(t(data.frame(yRepQ)))

names(yRepQdf) <- c(
  paste("L", 1:length(alphas), sep = ""),
  paste("U", 1:length(alphas), sep = "")
)


ppcDF <- cbind(
  sldf[order(sldf$fpHz),], yRepQdf
)


```




```{r}
ppc_freq <- gg_basic(ppcDF, aes(x = fpHz, group = Building, y = RpdB),
         ymin = 10, ymax = 80)+
  geom_point(aes(color = "Data"))+
  geom_line(aes(color = "Data"),alpha = 0.3)+
  geom_line(aes(y = L3, color = "99 p"), alpha = 0.1)+
  geom_line(aes(y = U3, color = "99 p"), alpha = 0.1)+
  geom_line(aes(y = L2, color = "95 p"), alpha = 0.1)+
  geom_line(aes(y = U2, color = "95 p"), alpha = 0.1)+
  geom_line(aes(y = L1, color = "50 p"), alpha = 0.1)+
  geom_line(aes(y = U1, color = "50 p"), alpha = 0.1)+
  scale_color_manual(
    values = c("green", "blue", "orange", "black"),
    guide = guide_legend(override.aes = list(alpha = 1)))+
  #scale_fill_manual(values = c("lightgreen","lightblue","grey"))+
    labs(color = "Predictive interval  ")
    #theme(legend.position = "bottom")

ppc_freq


```


```{r}
mat1 <- OutOfRange(ppcDF, "RpdB", "L1", "U1", result = "matrix")
mat2 <- OutOfRange(ppcDF, "RpdB", "L2", "U2", result = "matrix")
mat3 <- OutOfRange(ppcDF, "RpdB", "L3", "U3", result = "matrix")


expected <- 27*alphas


oorBias1 <- apply(mat1, 2, sum) - expected[1]
oorBias2 <- apply(mat2, 2, sum) - expected[2]
oorBias3 <- apply(mat3, 2, sum) - expected[3]

oorBias1sd <- oorBias1/sqrt(27*alphas[1]*(1-alphas[1]))
oorBias2sd <- oorBias2/sqrt(27*alphas[2]*(1-alphas[2]))
oorBias3sd <- oorBias3/sqrt(27*alphas[3]*(1-alphas[3]))


prfct <- function(ex, diff, prop){
   (1-pbinom(ex+diff,27,prop)+ pbinom(ex-diff,27,prop))
}


prob1 <- prfct(expected[1], diff = abs(oorBias1), alphas[1])
prob2 <- prfct(expected[2], diff = abs(oorBias2), alphas[2])
prob3 <- prfct(expected[3], diff = abs(oorBias3), alphas[3])



oorDf <- data.frame(
  "fpHz" = rep(freq, 3),
  "OORbias" = c(oorBias1,oorBias2,oorBias3),
  "OORbiasSd" = c(oorBias1sd,oorBias2sd,oorBias3sd),
  "alpha" = paste(rep(1-alphas, each = 21), "p"),
  "probs" = c(prob1,prob2,prob3)
)





ggplot(oorDf)+
  aes(x = fpHz, y = OORbias, color = alpha)+
  geom_point()+
  geom_line()+
  scale_x_continuous(trans = log2_trans(),
                       minor_breaks = c(freq,6300,8000,10000),
                       breaks = c(63,125,250,500,1000,2000,4000,8000),
                       labels = c(63,125,250,500,1000,2000,4000,8000),
    )+
  scale_color_manual(
    values = c("green", "blue", "orange"),
    breaks = c("0.5 p", "0.95 p", "0.99 p"))+
  labs(y = "OOR - Expected",
       x = expression(paste("Frequency ",
                             italic(" f"), " / Hz")),
       color = "Predictive interval")+
  theme(legend.position = "bottom")


ppcFreqSt <- ggplot(oorDf)+
  aes(x = fpHz, y = OORbiasSd, color = alpha)+
  geom_point()+
  geom_line()+
  scale_x_continuous(trans = log2_trans(),
                       minor_breaks = c(freq,6300,8000,10000),
                       breaks = c(63,125,250,500,1000,2000,4000,8000),
                       labels = c(63,125,250,500,1000,2000,4000,8000),
    )+
  scale_y_continuous(breaks = -3:3)+
  scale_color_manual(
    values = c("green", "blue", "orange"),
    breaks = c("0.5 p", "0.95 p", "0.99 p"))+
  labs(y = "(OOR - Expected) / Standard deviation",
       x = expression(paste("Frequency ",
                             italic(" f"), " / Hz")),
       color = "Predictive interval  ")+
  theme(legend.position = "bottom")+
  coord_cartesian(ylim = c(-3,3))

ppcFreqSt


ppcFreqBinom <- ggplot(oorDf)+
  aes(x = fpHz, y = probs, color = alpha)+
  geom_point()+
  geom_line(linetype = "dotted")+
  scale_x_continuous(trans = log2_trans(),
                       minor_breaks = c(freq,6300,8000,10000),
                       breaks = c(63,125,250,500,1000,2000,4000,8000),
                       labels = c(63,125,250,500,1000,2000,4000,8000),
    )+
  scale_color_manual(
    values = c("green", "blue", "orange"),
    breaks = c("0.5 p", "0.95 p", "0.99 p"))+
  labs(y = "P(|Binominal - Expected| >= |OOR-Expected|)",
       x = expression(paste("Frequency ",
                             italic(" f"), " / Hz")),
       color = "Predictive interval")+
  theme(legend.position = "bottom")

ppcFreqBinom


```





```{r}

oor <- c()

oor[1] <- OutOfRange(ppcDF, "RpdB", "L1", "U1", "sum")
oor[2] <- OutOfRange(ppcDF, "RpdB", "L2", "U2", "sum")
oor[3] <- OutOfRange(ppcDF, "RpdB", "L3", "U3", "sum")


data.frame(
  "Expected" = alphas*567,
  "Out of range" = oor
)



```






## PPC Whiteining TVAR Mahalanobis 


```{r}

# PPCwhiteSample <- apply(
#   exposter$y_rep,
#   1,
#   whiteningTVAR,
#   maxlag = 1
# )
# 
# PPCwhiteSample <- t(PPCwhiteSample)
# saveRDS(PPCwhiteSample, file = "Bayes/Saves/PPCwhite")

PPCwhiteSample <- readRDS("Bayes/Saves/PPCwhite")



```


```{r}
ntp_datamat <- matrix(sldf$RpdB, nrow = 27,byrow = T)

ntp_datawhite <- whiteningTVAR(ntp_datamat, 1)

```


```{r}


alphas <- c(0.5,0.05,0.01)

sim_quant <- apply(
  PPCwhiteSample, 2, quantile, probs = c(alphas/2, 1-alphas/2)
)

sim_quant <- data.frame(t(sim_quant))
names(sim_quant) <- c(
  paste("L", 1:length(alphas), sep = ""),
  paste("U", 1:length(alphas), sep = "")
)


own_ribbon <- data.frame(
  Ind = 1:27,
  ydata = ntp_datawhite,
  sim_mean = apply(PPCwhiteSample, 2, mean),
  sim_quant
)


basis <- ggplot(own_ribbon)+
    aes(x = Ind)+
    geom_ribbon(aes(ymin = L3, ymax = U3, fill = "99 p"), alpha = 0.1)+
    geom_ribbon(aes(ymin = L2, ymax = U2, fill = "95 p"), alpha = 0.2)+
    geom_ribbon(aes(ymin = L1, ymax = U1, fill = "50 p"), alpha = 0.3)+
    geom_line(aes(y = sim_mean, color = "PPC mean"))


ppcWhite <- basis +
    geom_point(aes(y = ydata, color = "Data"))+
    geom_line(aes(y= ydata, color = "Data"))+
    scale_fill_manual(values = rep("black", 3))+
    scale_color_manual(values = c("blue", "black"))+
    labs(color = "", fill = "Predictive interval",
         x = "Wall", y = "Squared mahalanobis")+
    theme(legend.position = "bottom")+
  scale_x_continuous(breaks = seq(3,27,3),minor_breaks = 1:27)

ppcWhite


```



```{r}

inrangeWhite <- c()

inrangeWhite[1] <- sum(own_ribbon$ydata >=own_ribbon$L1 & own_ribbon$ydata <=own_ribbon$U1)
inrangeWhite[2] <- sum(own_ribbon$ydata >=own_ribbon$L2 & own_ribbon$ydata <=own_ribbon$U2)
inrangeWhite[3] <- sum(own_ribbon$ydata >=own_ribbon$L3 & own_ribbon$ydata <=own_ribbon$U3)


oor <- 27-inrangeWhite


data.frame(
  "Predictive interval" = alphas*27,
  "Out of range" = oor,
  "Standardized OOR" = (oor - alphas*27)/sqrt(27*alphas*(1-alphas))
)



```



```{r}


mahalanobisOutlier <- gg_basic(sldf_fit[is.element(sldf_fit$Building,c(16,13,20,5) ),], aes(x = fpHz, group = Building), ymin = 20, ymax = 80)+
    geom_point(aes(y = RpdB, color = as.factor(Building)))+
    geom_line(aes(y = RpdB, color = as.factor(Building)), alpha = 1/3)+
    geom_line(aes(y = Mean, color = "Data mean"))+
  scale_color_manual(values = c(RavenR::ggplotColours(4), "black" ))+
  labs(color = "Wall")+
  theme(
    legend.position = c(0.7, 0.2),
    legend.justification = c(0.5, 0.5),
    legend.background = element_rect(fill = "transparent"),
    legend.spacing.y = unit(-0.7, "cm"),
    panel.grid.major = element_line(linewidth = 1))

mahalanobisOutlier

```





## PPC Rw




```{r}
alphas <- c(0.5,0.05,0.01)

#rwPPC <- apply(exposter$y_rep, 1:2, rw_value)
#saveRDS(rwPPC, file = "Bayes/Saves/PPCrw")

rwPPC <- readRDS("Bayes/Saves/PPCrw")

rwPPCquantiles <- apply(
  rwPPC, 2, quantile, probs = c(alphas/2, 1-alphas/2)
)


# hdi1 <- HDInterval::hdi(rwPPC, credMass  = 1-alphas[1])
# hdi2 <- HDInterval::hdi(rwPPC, credMass  = 1-alphas[2])
# hdi3 <- HDInterval::hdi(rwPPC, credMass  = 1-alphas[3])
# hdi <- rbind(hdi1,hdi2,hdi3)
# 
# rwPPCquantiles <- hdi[c(1,3,5,2,4,6),]

rwPPCquantiles <- data.frame(t(rwPPCquantiles))


names(rwPPCquantiles) <- c(
  paste("L", 1:length(alphas), sep = ""),
  paste("U", 1:length(alphas), sep = "")
)


rwData <- apply(sl_list_stan$y, 1, rw_value)


ownRibbonRw <- data.frame(
  Ind = 1:27,
  ydata = rwData,
  sim_mean = apply(rwPPCquantiles, 1, mean),
  rwPPCquantiles
)


```



```{r}

basis <- ggplot(ownRibbonRw)+
    aes(x = Ind)+
    geom_ribbon(aes(ymin = L3, ymax = U3, fill = "99 p"), alpha = 0.1)+
    geom_ribbon(aes(ymin = L2, ymax = U2, fill = "95 p"), alpha = 0.2)+
    geom_ribbon(aes(ymin = L1, ymax = U1, fill = "50 p"), alpha = 0.3)+
    geom_line(aes(y = sim_mean, color = "PPC mean"))


ppcRW <- basis +
    geom_point(aes(y = ydata, color = "Data"))+
    geom_line(aes(y= ydata, color = "Data"))+
    scale_fill_manual(values = rep("black", 3))+
    scale_color_manual(values = c("blue", "black"))+
    labs(color = "", fill = "Predictive interval",
         x = "Wall", y = bquote(Weighted~sound~reduction~index~R[w]))+
    theme(legend.position = "bottom")+
  scale_x_continuous(breaks = seq(3,27,3),minor_breaks = 1:27)

ppcRW

```



```{r}

inrangeRw <- c()

inrangeRw[1] <- sum(ownRibbonRw$ydata >=ownRibbonRw$L1 & ownRibbonRw$ydata <=ownRibbonRw$U1)
inrangeRw[2] <- sum(ownRibbonRw$ydata >=ownRibbonRw$L2 & ownRibbonRw$ydata <=ownRibbonRw$U2)
inrangeRw[3] <- sum(ownRibbonRw$ydata >=ownRibbonRw$L3 & ownRibbonRw$ydata <=ownRibbonRw$U3)


data.frame(
  "Predictive interval" = 27*alphas,
  "Out of range" = 27-inrangeRw
)



```





## Simulated example data




```{r}

set.seed(26)


public_data <- sldf


y <- rmvnorm(1,
             mean = f_model$fitted,
             sigma = kronecker(diag(27),f_model$cov_matrix)
             )

public_data$RpdB <- y[1,]

l1 <- runif(27, 3.2, 5.1)
l2 <- runif(27, 2.5, 16/l1)

public_data$s <- rep(l1*l2, each = 21)
public_data$u <- rep(2*(l1+l2), each = 21)


gg_basic(public_data, aes(x = fpHz, y = RpdB, group = Building))+
  geom_point()+
  geom_line(alpha = 1/3)


# write.csv2(public_data, "publicDataSRI.csv", row.names = F)



```






